Power simulation section

### 2023-09-07 docs/manuscript.Rmd
add/work on more comments

<!-- Some basics on statistical power -->
In empirical research relying on hypothesis testing, the most common strategy for determining an adequate sample size is based on statistical power <!-- FP:  cite lakens again here? -->. Statistical power is defined as the probability that a hypothesis test turns out significant <!-- FP: ... has a significant p-value? to make sure that we mean the statistical/technical meaning of significant  --> when analyzing repeated samples from a population with a true effect of some pre-specified size. Less formally, power is also described as the probability that a hypothesis test accepts the alternative hypothesis $H_1$ if $H_1$ is indeed true. If the sample size (i.e., the number of participants and/or stimuli) used for data collection is insufficient to reliably detect <!-- FP: alternative to reliably detect: detect the effects with high probability --> the effects or relationships being investigated, this particular study would be considered "under-powered".

<!-- what are the resulting issues with under-powered studies? -->
Conducting under-powered research has many negative consequences. First, relying on under-powered experiments may yield inconclusive (if low power is discussed by the researchers) or misleading (if low power is ignored) results, hindering the accumulation of knowledge. Second, under-powered studies waste resources by consuming time, effort, and funding without delivering meaningful results, thereby missing valuable discoveries and opportunities for scientific advancement <!-- FP: I would cut everything after funding  -->.

@@

<!-- advantages of an a priori power analysis -->
Conducting a careful a priori power analysis helps researchers decide which experimental design and statistical models are both feasible and appropriate for analyzing the data and answering their research questions.
Also, while conducting a proper power analysis, one has to consider every aspect of the experimental design and will possibly notice statistical or design challenges before the start of an experiment. 
This gives researchers a great opportunity to improve the design. Especially in fast-moving research fields, adding a solid sample size calculation into the process can act as a safeguard for ensuring high-quality research.

<!-- external incentives to do a power analysis -->
Moreover, many journals and funding agencies now mandate the inclusion of power analysis as part of study protocols and grant proposals, recognizing its significance in ensuring robust and meaningful findings. By incorporating power analysis into research planning, researchers can enhance the replicability and credibility of their work, contributing to the advancement of the respective research field through more robust insights. <!-- FP: this last sentence seems like a repetition  -->

<!-- standard software is available for "simple" models and study designs -->
For simple statistical models, like t-tests, ANOVA, and linear regression, with common study designs (e.g., mean comparison between two groups) user-friendly software for power analysis is readily available [@faulStatisticalPowerAnalyses2009].

@@

The inequality of conditional and marginal effects in GLMMs (i.e., how predictor variables affect the response within groups versus their overall impact across all groups, which can differ due to random effects) <!-- FP: I have tried to introduce contidional and marginal in the above section, so I would suggest to remove the bracket here --> [@fahrmeirRegressionModelsMethods2021] makes their interpretation more difficult. One must be careful when specifying parameter values based on previous studies or pilot data that use the marginal interpretation (e.g., a pilot study providing an estimate of how often neurologists make an accurate diagnosis based on brain scans). <!-- SG: What does that mean for someone who wants to do this then? FP: is it more clear now? --> However, this does not mean that we cannot use the marginal interpretation (average probability across persons and items) to inform plausible parameter values: When parameter values have been selected, we can compute the implied marginal distributions and compare this information to our domain knowledge. Then we can iteratively adjust the parameter values until we are satisfied with the implied distributions.

>>>>

<!-- Some basics on statistical power -->
In empirical research relying on hypothesis testing, the most common strategy for determining an adequate sample size is based on statistical power <!-- FP:  cite lakens again here? -->. Statistical power is defined as the probability that a hypothesis test turns out significant <!-- FP: ... probability that a hypothesis test has a significant p-value? to make sure that we mean the statistical/technical meaning of significant  --> when analyzing repeated samples from a population with a true effect of some pre-specified size. Less formally, power is also described as the probability that a hypothesis test accepts the alternative hypothesis $H_1$ if $H_1$ is indeed true. If the sample size (i.e., the number of participants and/or stimuli) used for data collection is insufficient to reliably detect <!-- FP: alternative to reliably detect: detect the effects with high probability --> the effects or relationships being investigated, this particular study would be considered "under-powered".

<!-- what are the resulting issues with under-powered studies? -->
Conducting under-powered research has many negative consequences. First, relying on under-powered experiments may yield inconclusive (if low power is taken into account) or misleading (if low power is ignored) results, hindering the accumulation of knowledge. Second, under-powered studies waste resources by consuming time, effort, and funding without delivering meaningful results, thereby missing valuable discoveries and opportunities for scientific advancement <!-- FP: I would cut everything after funding  -->.

@@

<!-- advantages of an a priori power analysis -->
Conducting a careful a priori power analysis helps researchers decide which experimental design and statistical models are both feasible and appropriate for analyzing the data and answering their research questions.
Also, while conducting a proper power analysis, one has to consider every aspect of the experimental design and will possibly notice statistical or design challenges before starting with data collection. 
This gives researchers a great opportunity to improve the design. Especially in fast-moving research fields, adding a solid sample size calculation to the process can act as a safeguard for ensuring high-quality research.

<!-- external incentives to do a power analysis -->
Moreover, many journals and funding agencies now require that a power analysis is included in study protocols and grant proposals, recognizing its significance in ensuring robust and meaningful findings. By incorporating power analysis into research planning, researchers can enhance the replicability and credibility of their work, contributing to the advancement of the respective research field through more robust insights. <!-- FP: this last sentence seems like a repetition. wouldn't some sentence better fit here stating that proficiency with power analysis improves your career chances in academia by increasing the probability to get papers published and grants funded (which is a new argument we haven't already made elsewhere)  -->

<!-- standard software is available for "simple" models and study designs -->
For simple statistical models, like t-tests, ANOVA, and linear regression, with common study designs (e.g., mean comparison between two groups) user-friendly software for power analysis is readily available [@faulStatisticalPowerAnalyses2009]. <!-- FP: we could reintroduce the old section her on: with more complex models and designs, power analysis is more complicated because no analytic solution is possible, power has to be estimated by simulation, and there is no user-friendly software that includes all relevant designs.  -->

@@

The inequality of conditional and marginal effects in GLMMs (i.e., how predictor variables affect the response within groups versus their overall impact across all groups, which can differ due to random effects) <!-- FP: I have tried to introduce conditional and marginal in the above section, so I would suggest to remove the bracket here --> [@fahrmeirRegressionModelsMethods2021] makes their interpretation more difficult. One must be careful when specifying parameter values based on previous studies or pilot data that use the marginal interpretation (e.g., a pilot study providing an estimate of how often neurologists make an accurate diagnosis based on brain scans). <!-- SG: What does that mean for someone who wants to do this then? FP: is it more clear now? --> However, this does not mean that we cannot use the marginal interpretation (average probability across persons and items) to inform plausible parameter values: When parameter values have been selected, we can compute the implied marginal distributions and compare this information to our domain knowledge. Then we can iteratively adjust the parameter values until we are satisfied with the implied distributions.
